---
title: "ggplots week 8 hw"
output: md_document
---


# Example Data: Housing prices:
```{r}

housing <- read.csv("~/Desktop/eeb-177/lab-work/exercise-8/Rgraphics/dataSets/landdata-states.csv")
head(housing[1:5])
```
# ggplot2 VS Base for simple graphs
```{r}
hist(housing$Home.Value)

```
# ggplot2 histogram example: 

```{r}
library(ggplot2)
ggplot(housing, aes(x = Home.Value)) +
  geom_histogram()

```

# ggplot2 Base graphics VS ggplot for more complex graphs:

```{r}
plot(Home.Value ~ Date,
     data=subset(housing, State == "MA"))
points(Home.Value ~ Date, col="red",
       data=subset(housing, State == "TX"))
legend(1975, 400000,
       c("MA", "TX"), title="State",
       col=c("black", "red"),
       pch=c(1, 1))
```

# ggplot2 colored scatter plot example: 

```{r}
ggplot(subset(housing, State %in% c("MA", "TX")),
       aes(x=Date,
           y=Home.Value,
           color=State))+
  geom_point()
```

# Geometic Objects (geom)

```{r}
help.search("geom_", package = "ggplot2")

```

# Points (Scatterplot):

```{r}
hp2001Q1 <- subset(housing, Date == 2001.25) 
ggplot(hp2001Q1,
       aes(y = Structure.Cost, x = Land.Value)) +
  geom_point()

```

```{r}
ggplot(hp2001Q1,
       aes(y = Structure.Cost, x = log(Land.Value))) +
  geom_point()
```

# Lines (Prediction Line): 

```{r}
hp2001Q1$pred.SC <- predict(lm(Structure.Cost ~ log(Land.Value), data = hp2001Q1))

p1 <- ggplot(hp2001Q1, aes(x = log(Land.Value), y = Structure.Cost))

p1 + geom_point(aes(color = Home.Value)) +
  geom_line(aes(y = pred.SC))

```

# Smoothers

```{r}
p1 +
  geom_point(aes(color = Home.Value)) +
  geom_smooth()


```


# Text (Label Points):

```{r}
p1 + 
  geom_text(aes(label=State), size = 3)

```

```{r}
library("ggrepel")
p1 + 
  geom_point() + 
  geom_text_repel(aes(label=State), size = 3)

```

# Aesthetic Mapping VS Assignment:

```{r}
 p1 +
    geom_point(aes(size = 2),
               color="red") 
```


# Mapping Variables To Other Aesthetics

```{r}
p1 +
  geom_point(aes(color=Home.Value, shape = region))

```

# Exercise I:

```{r}
library(ggplot2)

dat <- read.csv("~/Desktop/eeb-177/lab-work/exercise-8/Rgraphics/dataSets/EconomistData.csv")
head(dat)

ggplot(dat, aes(x = CPI, y = HDI, size = HDI.Rank)) + geom_point()

```
# 1. Create a scatter plot with CPI on the x axis and HDI on the y axis.
# 2. Color the points blue.
# 3. Map the color of the the points to Region.
# 4. Make the points bigger by setting size to 2
# 5. Map the size of the points to HDI.Rank

```{r}
library(ggplot2)

ggplot(dat, aes(x = CPI, y = HDI)) + geom_point(color="blue")
```

```{r}
ggplot(dat, aes(x = CPI, y = HDI, color = Region)) + geom_point()
```

```{r}
ggplot(dat, aes(x = CPI, y = HDI, color = Region)) + geom_point(size=2)
```

```{r}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point(aes(color = Region, size =  HDI.Rank))
```

# Statistical Transformations:

```{r}
 args(geom_histogram)
  args(stat_bin)
```

# Setting Statistical Transformation Arguments:

```{r}
housing <- read.csv("~/Desktop/eeb-177/lab-work/exercise-8/Rgraphics/dataSets/landdata-states.csv")

p2 <- ggplot(housing, aes(x = Home.Value))
p2 + geom_histogram()

```

```{r}
  p2 + geom_histogram(stat = "bin", binwidth=4000)
```

# Changing The Statistical Transformation:

```{r}
housing.sum <- aggregate(housing["Home.Value"], housing["State"], FUN=mean)
rbind(head(housing.sum), tail(housing.sum))
```

```{r}
ggplot(housing.sum, aes(x=State, y=Home.Value)) + 
  geom_bar(stat="identity")
```

# Exercise II:

# 1. Re-create a scatter plot with CPI on the x axis and HDI on the y axis (as you did in the previous exercise).
# 2. Overlay a smoothing line on top of the scatter plot using geom_smooth, but use a linear model for the predictions. Hint: see ?stat_smooth.
# 3. Overlay a smoothing line on top of the scatter plot using geom_line. Hint: change the statistical transformation.
# 4. BONUS: Overlay a smoothing line on top of the scatter plot using the default loess method, but make it less smooth. Hint: see ?loess.

```{r}
 ggplot(dat, aes(x = CPI, y = HDI)) +
    geom_point()
```

```{r}
 ggplot(dat, aes(x = CPI, y = HDI)) +
    geom_point() +
    geom_smooth(method = "lm")
```

```{r}
 ggplot(dat, aes(x = CPI, y = HDI)) +
    geom_point() +
    geom_smooth()
```

```{r}
 ggplot(dat, aes(x = CPI, y = HDI)) +
    geom_point() +
    geom_smooth(span = .4)
```

# Scale Modification Examples:

```{r}
 p3 <- ggplot(housing, aes(x = State, y = Home.Price.Index)) + theme(legend.position="top", axis.text=element_text(size = 6))
(p4 <- p3 + geom_point(aes(color = Date),
                       alpha = 0.5,
                       size = 1.5,
                       position = position_jitter(width = 0.25, height = 0)))
```

# Scale Modification Examples:

```{r}
 p3 <- ggplot(housing,
               aes(x = State,
                   y = Home.Price.Index)) + 
          theme(legend.position="top",
                axis.text=element_text(size = 6))
  (p4 <- p3 + geom_point(aes(color = Date),
                         alpha = 0.5,
                         size = 1.5,
                         position = position_jitter(width = 0.25, height = 0)))
```
```{r}
p4 + scale_x_discrete(name="State Abbreviation") +
    scale_color_continuous(name="",
                           breaks = c(19751, 19941, 20131),
                           labels = c(1971, 1994, 2013))
```

```{r}
  p4 +
    scale_x_discrete(name="State Abbreviation") +
    scale_color_continuous(name="",
                           breaks = c(19751, 19941, 20131),
                           labels = c(1971, 1994, 2013),
                           low = "blue", high = "red")
```

```{r}
p4 + scale_x_discrete(name="State Abbreviation") +  scale_color_continuous(name="", breaks = c(1976, 1994, 2013), labels = c("'76", "'94", "'13"))
```

# Using different color scales:
```{r}
p4 + scale_x_discrete(name="State Abbreviation") + scale_color_continuous(name="", breaks = c(1976, 1994, 2013), labels = c("'76", "'94", "'13"), low = "blue", high = "red")

```

# Exercise III:

# 1: Create a scatter plot with CPI on the x axis and HDI on the y axis. Color the points to indicate region.
# 2: Modify the x, y, and color scales so that they have more easily-understood names (e.g., spell out “Human development Index” instead of “HDI”).
# 3: Modify the color scale to use specific values of your choosing. Hint: see ?scale_color_manual.


```{r}
  ggplot(dat, aes(x = CPI, y = HDI, color = "Region")) +
    geom_point()
```

```{r}
ggplot(dat, aes(x = CPI, y = HDI, color = "Region")) +
  geom_point() +
  scale_x_continuous(name = "Corruption Perception Index") +
  scale_y_continuous(name = "Human Development Index") +
  scale_color_discrete(name = "Region of the world")

```

```{r}
ggplot(dat, aes(x = CPI, y = HDI, color = "Region")) +
  geom_point() +
  scale_x_continuous(name = "Corruption Perception Index") +
  scale_y_continuous(name = "Human Development Index") +
    scale_color_manual(name = "Region of the world",
                       values = c("#24576D",
                                  "#099DD7",
                                  "#28AADC",
                                  "#248E84",
                                  "#F2583F",
                                  "#96503F"))
```

# Faceting
# What is the trend in housing prices in each state?

```{r}
  p5 <- ggplot(housing, aes(x = Date, y = Home.Value))
  p5 + geom_line(aes(color = State))  
```

# Faceting to the rescue:

```{r}
 (p5 <- p5 + geom_line() +
     facet_wrap(~State, ncol = 10))
```

# Themes:

```{r}
 p5 + theme_linedraw()
```

```{r}
 p5 + theme_light()
```

# Overriding theme defaults:

```{r}
  p5 + theme_minimal() +
    theme(text = element_text(color = "turquoise"))
```

# Creating and saving new themes:

```{r}
  p5 + theme_minimal() +
    theme(text = element_text(color = "turquoise"))

```
```{r}
theme_new <- theme_bw() +
    theme(plot.background = element_rect(size = 1, color = "blue", fill = "black"),
          text=element_text(size = 12, family = "Serif", color = "ivory"),
          axis.text.y = element_text(colour = "purple"),
          axis.text.x = element_text(colour = "red"),
          panel.background = element_rect(fill = "pink"),
          strip.background = element_rect(fill ="orange"))

  p5 + theme_new

```

# The #1 FAQ:

```{r}
housing.byyear <- aggregate(cbind(Home.Value, Land.Value) ~ Date, data = housing, mean)
  ggplot(housing.byyear,
         aes(x=Date)) +
    geom_line(aes(y=Home.Value), color="red") +
    geom_line(aes(y=Land.Value), color="blue")

```
```{r}
  library(tidyr)
  home.land.byyear <- gather(housing.byyear,
                             value = "value",
                             key = "type",
                             Home.Value, Land.Value)
  ggplot(home.land.byyear,
         aes(x=Date,
             y=value,
             color=type)) +
    geom_line()

```

# Putting It All Together
# Challenge: Recreate This Economist Graph

```{r}
 dat <- read.csv("~/Desktop/eeb-177/lab-work/exercise-8/Rgraphics/dataSets/EconomistData.csv")

  pc1 <- ggplot(dat, aes(x = CPI, y = HDI, color = Region))
  pc1 + geom_point()
```
# Adding the trend line:

```{r}
  (pc2 <- pc1 +
     geom_smooth(aes(group = 1),
                 method = "lm",
                 formula = y ~ log(x),
                 se = FALSE,
                 color = "red")) +
     geom_point()
```

# Use open points:

```{r}
df2 <- data.frame(x = 1:5 , y = 1:25, z = 1:25)
  s <- ggplot(df2, aes(x = x, y = y))
  s + geom_point(aes(shape = z), size = 4) + scale_shape_identity()
  
  s + geom_point(aes(shape = z), size = 4, colour = "Red") +
    scale_shape_identity()
  s + geom_point(aes(shape = z), size = 4, colour = "Red", fill = "Black") +
    scale_shape_identity()
```

```{r}
 pc2 +
    geom_point(shape = 1, size = 4)
```


```{r}
  (pc3 <- pc2 +
     geom_point(size = 4.5, shape = 1) +
     geom_point(size = 4, shape = 1) +
     geom_point(size = 3.5, shape = 1))
```

# Labelling points:

```{r}
 pointsToLabel <- c("Russia", "Venezuela", "Iraq", "Myanmar", "Sudan",
                     "Afghanistan", "Congo", "Greece", "Argentina", "Brazil",
                     "India", "Italy", "China", "South Africa", "Spane",
                     "Botswana", "Cape Verde", "Bhutan", "Rwanda", "France",
                     "United States", "Germany", "Britain", "Barbados", "Norway", "Japan",
                     "New Zealand", "Singapore")
```


```{r}
(pc4 <- pc3 +
    geom_text(aes(label = Country),
              color = "gray20",
              data = subset(dat, Country %in% pointsToLabel)))
```


```{r}
 library("ggrepel")
  pc3 +
    geom_text_repel(aes(label = Country),
              color = "gray20",
              data = subset(dat, Country %in% pointsToLabel),
              force = 10)
```


# Change the region labels and order:

```{r}
  dat$Region <- factor(dat$Region,
                       levels = c("EU W. Europe",
                                  "Americas",
                                  "Asia Pacific",
                                  "East EU Cemt Asia",
                                  "MENA",
                                  "SSA"),
                       labels = c("OECD",
                                  "Americas",
                                  "Asia &\nOceania",
                                  "Central &\nEastern Europe",
                                  "Middle East &\nnorth Africa",
                                  "Sub-Saharan\nAfrica"))

```


```{r}
pc4$data <- dat
  pc4

```


# Add title and format axes:

```{r}
  library(grid)
  (pc5 <- pc4 +
    scale_x_continuous(name = "Corruption Perceptions Index, 2011 (10=least corrupt)",
                       limits = c(.9, 10.5),
                       breaks = 1:10) +
    scale_y_continuous(name = "Human Development Index, 2011 (1=Best)",
                       limits = c(0.2, 1.0),
                       breaks = seq(0.2, 1.0, by = 0.1)) +
    scale_color_manual(name = "",
                       values = c("#24576D",
                                  "#099DD7",
                                  "#28AADC",
                                  "#248E84",
                                  "#F2583F",
                                  "#96503F")) +
    ggtitle("Corruption and Human development"))
```



# Theme tweaks: 

```{r}
library(grid) # for the 'unit' function
  (pc6 <- pc5 +
    theme_minimal() + # start with a minimal theme and add what we need
    theme(text = element_text(color = "gray20"),
          legend.position = c("top"), # position the legend in the upper left 
          legend.direction = "horizontal",
          legend.justification = 0.1, # anchor point for legend.position.
          legend.text = element_text(size = 11, color = "gray10"),
          axis.text = element_text(face = "italic"),
          axis.title.x = element_text(vjust = -1), # move title away from axis
          axis.title.y = element_text(vjust = 2), # move away for axis
          axis.ticks.y = element_blank(), # element_blank() is how we remove elements
          axis.line = element_line(color = "gray40", size = 0.5),
          axis.line.y = element_blank(),
          panel.grid.major = element_line(color = "gray50", size = 0.5),
          panel.grid.major.x = element_blank()
          ))

```

# Add model R2 and source note:

```{r}
(mR2 <- summary(lm(HDI ~ log(CPI), data = dat))$r.squared)
```

```{r}
  library(grid)
 # png(file = "~/Desktop/eeb-177/lab-work/exercise-8/Rgraphics/images/econScatter10.png", width = 800, height = 600)
  pc6 
  grid.text("Sources: Transparency International; UN Human Development Report",
           x = .02, y = .03,
           just = "left",
           draw = TRUE)
  grid.segments(x0 = 0.81, x1 = 0.825,
                y0 = 0.90, y1 = 0.90,
                gp = gpar(col = "red"),
                draw = TRUE)
  grid.text(paste0("RÂ² =  0.5212859",
                   as.integer(mR2*100),
                   "%"),
            x = 0.835, y = 0.90,
            gp = gpar(col = "gray20"),
            draw = TRUE,
            just = "left")
  
 # dev.off()
```

